Clonal interference between aneuploidy and mutation

### 2021-05-24 ms/ms.tex
add todo for citation
continue with multiple-locus model

\paragraph*{Multiple-locus model.} It expands the \emph{single-mutant model} by allowing for any number of beneficial mutations.

>>>>

\paragraph*{Multiple-locus model.} It expands the \emph{single-mutant model} by allowing for any number of beneficial mutations. For each mutation $i$, selection coefficient $s_i$ is independently drawn from an exponential distribution with expected value $s$. The parameters of the model are $\mu$, the mutation rate; $\alpha$, the aneuploidy gain rate; $\alpha_l$, the aneuploidy loss rate; $s$, the parameter of the exponential distribution; $s_b$, fitness benefit of aneuploidy; $c$, fitness cost of aneuploidy, where $0<c<1$. In this model, the population may include any number of genotypes that contain zero or more mutations and zero or one extra chromosomes. The fitness $w_g$ of genotype $g$ is:

\begin{equation} 
w_g= 
\begin{cases}
   1, & \text{wildtype} \\
	1-c+s_b, & \text{aneuploid wildtype} \\
	\prod_{\text{i is mutation in g}}{(1+s_i)}, & \text{euploid mutant} \\ 
	s_b+(1-c)\prod_{\text{i is mutation in g}}{(1+s_i)}, & \text{aneuploid mutant}
\end{cases}
\end{equation}

Therefore, aneuploidy loss would be favored by selection only if there are enough beneficial mutations or/and the mutation selection coefficients $s_i$ are large enough. The intuition is as follows: when the benefit of the accumulated beneficial mutations is small, then the benefit of aneuploidy has a large effect; when the benefit of the accumulated beneficial mutations benefit is large, then aneuploidy doesn’t add much, and its cost becomes significant.

Unlike the single-locus model, the population size for the multiple-locus model is not constant. Indeed, a serial-transfer protocol~\citep{Barrick2013} is simulated, where the population is repeatedly diluted by transfer to a fresh medium, starting a new growth cycle, similar to the original experiment. In the model the population initial size is $N$, then it is doubled every iteration till reaching $8\cdot{N}$, then it is diluted back to $N$.

Denote by $n_g$ be the number of cells of genotype $g$ in the population. At each generation, genotypes frequencies are changed by the following steps:
\begin{enumerate}
	\item Selection step: 
	\begin{equation} \label{eq:multiple-selection} 
		n_g\gets{n_g\cdot{w_g}}
	\end{equation}
	\item Random drift step:
	\begin{equation} \label{eq:multiple-rand-drift} \begin{aligned}
			f \sim Mult(N', (\frac{n_1}{N'},\frac{n_2}{N'},...)) \text{ where } N'=n_1+n_2+... 
		\end{aligned}
	\end{equation}
	\item Normalization after N change (growth or dilution):
	\begin{equation} \label{eq:multiple-pop-size} 
		n_g\gets{f_g\cdot{N}}
	\end{equation}
\end{enumerate}
